Non-word repetition in children learning Yélî Dnye

### 2021-11-25 manuscript.tex
resubmitted

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  american,
  ,man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Non-word repetition in children learning Yélî Dnye},
  pdfauthor={Alejandrina Cristia1 \& Marisa Casillas2,3},
  pdflang={en-US},
  pdfkeywords={phonology, non-word repetition, Papuan, non-industrial, non-urban, comparative, typology, markedness, literacy},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{fontspec}

@@

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\else
  \usepackage[main=american]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Non-word repetition in children learning Yélî Dnye}
\author{Alejandrina Cristia\textsuperscript{1} \& Marisa Casillas\textsuperscript{2,3}}
\date{}

@@

Our focus on age stems from previous work, where performance increases with child age (Farmani et al., 2018; Kalnak, Peyrard-Janvid, Forssberg, \& Sahlén, 2014; Vance, Stackhouse, \& Wells, 2005). Although past research has not investigated potential correlations with birth order on NWR, there is a sizable literature on these correlations in other language tasks (e.g., Havron et al., 2019), and therefore we report on these too. Common explanations for advantages for first- over later-born children include differential allocation of familial resources, particularly parental behaviors of cognitive stimulation (Lehmann, Nuevo-Chiquero, \& Vidal-Fernandez, 2018). Regarding child sex, no significant correlation has been found in previous NWR research (Chiat \& Roy, 2007), and in other language tasks evidence is mixed. Finally, prior research using NWR varies on whether significant differences as a function of maternal education are reported. For instance, no significant difference was found some studies (Balladares, Marshall, \& Griffiths, 2016; Farmani et al., 2018; Kalnak, Peyrard-Janvid, Forssberg, \& Sahlén, 2014; Meir \& Armon-Lotem, 2017); whereas significant differences were reported in others (Santos, Frau, Labrevoit, \& Zebib, 2020; Tuller et al., 2018). In other lines of work, maternal education often correlates with child language outcomes, including vocabulary reports (Frank, Braginsky, Yurovsky, \& Marchman, 2017) and word comprehension studies (Scaff, 2019). The causal pathways explaining this correlation are complex, but one explanation that is often discussed involves more educated mothers talking more to their children (see discussion in Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020).

@@

The fourth research goal we pursued is to use NWR with non-Western, non-urban populations, speaking a language with a moderate to large phonological inventory (see Maddieson, 2005 for a broad classification of languages based on inventory size). Indeed, NWR has seldom been used outside of urban settings in Europe and North America (Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020; with exceptions including Gallagher, 2014). To our knowledge, it has never been used with speakers of languages having large phonological inventories (e.g., more than 34 consonants and 7 vowel qualities Maddieson, 2013b, 2013a).

There are no theoretical reasons to presume that the technique will not generalize to these new conditions. That said, Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020) recently reported relatively lower NWR scores among the Tsimane', a non-Western rural population, interpreting these findings as consistent with the hypothesis that lower levels of infant-directed speech and/or low prevalence of literacy in a population could lead to population-level differences in NWR scores.

In view of these results, it is important to bear in mind that NWR is a task developed in countries where literacy is widespread, and it is considered an excellent predictor of reading; for instance, better than rhyme awareness (e.g., Gathercole, Willis, \& Baddeley, 1991). Therefore, it may not be a general index of phonological development, but instead reflect certain non-universal language skills. Indeed, Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020) present their task as being a good index of the development of ``short-hand-like'' representations specifically, which could thus miss, for example, more holistic phonological and phonetic representations. We return to the question of what was measured here in the Discussion.

Aside from Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020)'s hypotheses just mentioned, we have found little discussion of linguistic differences (i.e., potential differences in NWR as a function of which specific language children are learning, and/or its typology) or cultural differences (i.e., potential differences in NWR as a function of other differences across human populations).\footnote{Please note that the linguistic and cultural differences discussed here are different from the differences discussed in the extensive literature on NWR by bilingual participants. In that literature, authors are concerned with individual variation in exposure to one (as opposed to other) languages among multilingual children, as variation in relative language experiences could mask potential effects of language impairment. To try to measure language abilities above and beyond relative levels of experience with a given language, authors have tried to build non-words that tap language-dependent or language-independent knowledge. For instance, Tuller et al. (2018) employed a set of non-words judged to be language independent and two others that were more aligned with either French or German. The intuition is that NWR will correlate with the relative levels of exposure to that language more strongly when items are aligned with a specific language (``language-dependent'') than when they are ``language-independent.'' To make this more precise, among bilingual children, those that have more experience with English than Spanish should perform better on English non-words than their peers with less English experience. Preliminary results of an ongoing meta-analysis suggest significant associations between exposure to a given language and performance in both language-dependent and language-independent NWR (Farabolini, Taboh, Ceravolo, \& Guerra, 2021). In any case, this line of research focuses on links between exposure to a given language and NWR performance. In contrast, when we discuss linguistic or cultural differences here, we ask the question of whether children vary in their performance as a function of which language they are learning (e.g., the language's typological properties) and/or their overall, absolute levels of language experience (not relative levels in a multilingual setting).}

@@

Most Yélî parents are swidden horticulturalists, who occasionally fish. Within a group of households, it is often the case that older adolescents and adults spend their day tending to their farm plots (which may not be nearby), bringing up water from the river, washing clothes, preparing food, and engaging in other such activities. Starting around age two years, children more often spend large swaths of their day playing, swimming, and foraging for fruit, nuts, and shellfish in large (\textasciitilde10 members) independent and mixed-age child play groups (Brown \& Casillas, in press; Casillas, Brown, \& Levinson, 2021). Formal education is a priority for Yélî families, and many young parents have themselves pursued additional education beyond what is locally available (Casillas, Brown, \& Levinson, 2021). Local schools are well out of walking distance for many children (i.e., more than 1 hour on foot or by canoe each day), so it is very common for households situated close to a school to host their school-aged relatives during the weekdays for long segments of the school year. Children start school often at around age seven, although the precise age depends on the child's readiness, as judged by their teacher.

@@

We add some observations that will help us integrate this study into the broader investigation of NWR across cultures. As mentioned previously, there is one report of relatively low NWR scores among the Tsimane', which the authors of that paper interpret as consistent with long-term effects of low levels of infant-directed speech (Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020). However, Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020) also point out that this is based on between-paper comparisons, and thus methods and myriad other factors have not been controlled for. The Yélî community can help us gain new insights into this matter because direct speech to children under 3 years is comparably infrequent in this community (in fact it may be infrequent in many settings, including urban ones Bunce et al., under review). Our sample also shares other societal characteristics wih the Tsimane' (e.g., the community is rural and relies on farming, children grow up in wide familial networks, Casillas, Brown, \& Levinson, 2021). Although infant-directed speech has been measured in different ways among the Tsimane' and the Yélî communities, our most comparable estimates at present suggest that Tsimane' young children are spoken to about 4.2 minutes per hour (Scaff, Stieglitz, Casillas, \& Cristia, under review), and Yélî children about 3.6 minutes per hour (Casillas, Brown, \& Levinson, 2021). Thus, if these input quantities in early childhood relate to lower NWR scores later in life, we should observe similarly low NWR scores here as in Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020).

@@

Many NWR studies are based on a fixed list of 12-16 items that vary in length between 1 and 4 syllables, often additionally varying syllable complexity and/or cluster presence and complexity, and always meeting the condition that they do not mean anything in the target language (e.g., Balladares, Marshall, \& Griffiths, 2016; Wilsenach, 2013). We kept the same variation in item length and requirement for not being meaningful in the language, but we did not vary syllable complexity or clusters because these are vanishingly rare in Yélî Dnye. We also increased the number of items an individual child would be tested on, such that a child would get up to 23 items to repeat (other work has also used up to 24-46 items: Jaber-Awida, 2018; Kalnak, Peyrard-Janvid, Forssberg, \& Sahlén, 2014; Piazzalunga, Previtali, Pozzoli, Scarponi, \& Schindler, 2019), with the entire test inventory of 40 final items distributed across children. We used a relatively large number of items to explore correlations with length and phonological complexity. However, aware that this large item inventory might render the task longer and more tiresome, we split items across children. Naturally, designing the task in this way may make the study of individual variation within the population more difficult because different children are exposed to different items.

@@

Additionally, we estimated the usage frequency of the phones present in the target items in a corpus of child-centered recordings (Casillas, Brown, \& Levinson, 2021). That corpus was constituted by sampling from audio-recordings (7--9 hours long), collected as 10 children aged between 1 month and 3 years went about their day. The researchers selected 9 2.5-minute clips randomly and 11 1- or 5-minute clips by hand (selected to represent peak turn-taking and child vocal activity). These clips were segmented and transcribed by the lead researcher and a highly knowledgeable local assistant, who speaks Yélî Dnye natively, has ample experience in this kind of research, and often knew all the recorded people personally. For more details, please refer to Casillas, Brown, and Levinson (2021).

@@

Previous work typically reports two scores: a binary word-level exact repetition score, and a phoneme-level score, defined as the number of phonemes that can be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt; as in Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020). Previous work does not use distance metrics, but we report these rather than the phoneme-level scores because they are more informative. To illustrate these scores, recall our example of an English target being /bilik/ with an imagined response {[}bilig{]}. We would score this response as follows: at the whole item level this production would receive a score of zero (because the repetition is not exact); at the phoneme level this production would receive a score of 80\% (4 out of 5 phonemes repeated exactly); and the phone-based Levenshtein distance for this production is 20\% (because 20\% of phonemes were substituted or deleted). Notice that the phone-based Levenshtein distance is the complement of the phoneme-level NWR score. An advantage of using phone-based Levenshtein distance is that it is scored automatically with a script, and it can then easily be split in terms of deletions and substitutions (insertions were not attested in this study).

@@

There are some reasons to believe that Yélî Dnye put that hypothesis to a critical test: The phoneme inventory is both large and acoustically packed, in addition to containing several typologically infrequent (or unique) contrasts. One could then predict that correlations with typological frequency should be relatively weak because the ambient language puts more pressure on Yélî children to distinguish (perceptually and articulatorily) fine-grained phonetic differences than what is required of child speakers of other languages. On the other hand, it is also possible that this pressure gives Yélî children no benefit, and that some of these categories are simply acquired later in development. We can draw a parallel with children learning another Papuan language, Ku Waru, which has a packed inventory of lateral consonants; children do not produce adult-like realizations of the more complex of these laterals (the pre-stopped velar lateral /ɡ͡ʟ/) until 5 or 6 years of age (Rumsey, 2017)

@@

We investigated the effect of item complexity on NWR scores by varying the number of syllables in the item. In broad terms, children should have higher NWR scores for shorter items. That said, previous work summarized in the Introduction has shown both very small (e.g., Piazzalunga, Previtali, Pozzoli, Scarponi, \& Schindler, 2019) and very large (e.g., Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020) effects of stimulus length. Setting aside our monosyllabic stimuli (which contained typologically infrequent segments with lower NWR scores, as just discussed), we examined effects of item length among the remaining stimuli, which range between 2 and 4 syllables long. The effect of item length was not significant in a statistical model that additionally accounted for age and random effects of item and participant. We do not have a good explanation for why samples in the literature vary so much in terms of the size of length effects, but two possibilities are that this is not truly a length effect but a confound with some other aspect of the stimuli, or that there is variation in phonological representations that is poorly understood. We explain each idea in turn.

@@

In broad terms, we expected that NWR scores would increase with participant age, as this is the pattern observed in several previous studies (English Vance, Stackhouse, \& Wells, 2005; Italian Piazzalunga, Previtali, Pozzoli, Scarponi, \& Schindler, 2019; Cantonese Stokes, Wong, Fletcher, \& Leonard, 2006; but not in Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020). Indeed, age was significantly correlated with NWR scores and it also showed up as a significant predictor of NWR score when included as a control factor in the analyses of both item length and average segmental frequency. In brief, our results underscore the idea that phonological development continues well past the first few years of life, extending into middle childhood and perhaps later (Hazan \& Barrett, 2000; Rumsey, 2017).

In contrast, previous work varies with respect to correlations of NWR scores with maternal education (e.g., Farmani et al., 2018; Kalnak, Peyrard-Janvid, Forssberg, \& Sahlén, 2014; Meir \& Armon-Lotem, 2017). We did not expect large correlations with maternal education in our sample for two reasons: First, education on Rossel Island is generally highly valued and so widespread that little variation is seen there; second, formal education is not at all essential to ensuring one's success in society and may not be a reliable index of local socioeconomic variation locally. In fact, maternal education correlated with NWR score at about r\textasciitilde.1, which is small. We find correlations of about that size for participant sex, which is aligned with previous work (Chiat \& Roy, 2007).

@@

The fourth research area to which we wanted to contribute pertained to the use of NWR across languages and populations, since when designing this study we wondered whether NWR was a \emph{culture-fair} test of phonological development. Although our data cannot answer this question because we have only sampled one language and population here, we would like to spend some time discussing the integration of these results to the wider NWR literature. It is important to note at the outset that we cannot obtain a final answer because integration across studies implies not only variation in languages and child-rearing settings, but also in methodological aspects including non-word length, non-word design (e.g., the syllable and phone complexity included in the items), and task administration, among others. Nonetheless, we feel the NWR task is prevalent enough to warrant discussion about this, similarly to other tasks sometimes used to describe and compare children's language skills across populations, like the recent re-use of the MacArthur-Bates Communicative Development Inventory to look at vocabulary acquisition across multiple languages (Frank, Braginsky, Yurovsky, \& Marchman, 2017).

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  american,
  ,man,floatsintext]{apa6}
\title{Non-word repetition in children learning Yélî Dnye}
\author{Alejandrina Cristia\textsuperscript{1} \& Marisa Casillas\textsuperscript{2,3}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Non-word repetition in children learning Yélî Dnye},
  pdfauthor={Alejandrina Cristia1 \& Marisa Casillas2,3},
  pdflang={en-US},
  pdfkeywords={phonology, non-word repetition, Papuan, non-industrial, non-urban, comparative, typology, markedness, literacy},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{fontspec}

@@

\linenumbers
\usepackage{csquotes}
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\else
  \usepackage[main=american]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

Our focus on age stems from previous work, where performance increases with child age (Farmani et al., 2018; Kalnak, Peyrard-Janvid, Forssberg, \& Sahlén, 2014; Vance, Stackhouse, \& Wells, 2005). Although past research has not investigated potential correlations with birth order on NWR, there is a sizable literature on these correlations in other language tasks (e.g., Havron et al., 2019), and therefore we report on these too. Common explanations for advantages for first- over later-born children include differential allocation of familial resources, particularly parental behaviors of cognitive stimulation (Lehmann, Nuevo-Chiquero, \& Vidal-Fernandez, 2018). Regarding child sex, no significant correlation has been found in previous NWR research (Chiat \& Roy, 2007), and in other language tasks evidence is mixed. Finally, prior research using NWR varies on whether significant differences as a function of maternal education are reported. For instance, no significant difference was found some studies (Balladares, Marshall, \& Griffiths, 2016; Farmani et al., 2018; Kalnak et al., 2014; Meir \& Armon-Lotem, 2017); whereas significant differences were reported in others (Santos et al., 2020; Tuller et al., 2018). In other lines of work, maternal education often correlates with child language outcomes, including vocabulary reports (Frank, Braginsky, Yurovsky, \& Marchman, 2017) and word comprehension studies (Scaff, 2019). The causal pathways explaining this correlation are complex, but one explanation that is often discussed involves more educated mothers talking more to their children (see discussion in Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020).

@@

The fourth research goal we pursued is to use NWR with non-Western, non-urban populations, speaking a language with a moderate to large phonological inventory (see Maddieson, 2005 for a broad classification of languages based on inventory size). Indeed, NWR has seldom been used outside of urban settings in Europe and North America (Cristia et al., 2020; with exceptions including Gallagher, 2014). To our knowledge, it has never been used with speakers of languages having large phonological inventories (e.g., more than 34 consonants and 7 vowel qualities Maddieson, 2013b, 2013a).

There are no theoretical reasons to presume that the technique will not generalize to these new conditions. That said, Cristia et al. (2020) recently reported relatively lower NWR scores among the Tsimane', a non-Western rural population, interpreting these findings as consistent with the hypothesis that lower levels of infant-directed speech and/or low prevalence of literacy in a population could lead to population-level differences in NWR scores.

In view of these results, it is important to bear in mind that NWR is a task developed in countries where literacy is widespread, and it is considered an excellent predictor of reading; for instance, better than rhyme awareness (e.g., Gathercole, Willis, \& Baddeley, 1991). Therefore, it may not be a general index of phonological development, but instead reflect certain non-universal language skills. Indeed, Cristia et al. (2020) present their task as being a good index of the development of ``short-hand-like'' representations specifically, which could thus miss, for example, more holistic phonological and phonetic representations. We return to the question of what was measured here in the Discussion.

Aside from Cristia et al. (2020)'s hypotheses just mentioned, we have found little discussion of linguistic differences (i.e., potential differences in NWR as a function of which specific language children are learning, and/or its typology) or cultural differences (i.e., potential differences in NWR as a function of other differences across human populations).\footnote{Please note that the linguistic and cultural differences discussed here are different from the differences discussed in the extensive literature on NWR by bilingual participants. In that literature, authors are concerned with individual variation in exposure to one (as opposed to other) languages among multilingual children, as variation in relative language experiences could mask potential effects of language impairment. To try to measure language abilities above and beyond relative levels of experience with a given language, authors have tried to build non-words that tap language-dependent or language-independent knowledge. For instance, Tuller et al. (2018) employed a set of non-words judged to be language independent and two others that were more aligned with either French or German. The intuition is that NWR will correlate with the relative levels of exposure to that language more strongly when items are aligned with a specific language (``language-dependent'') than when they are ``language-independent.'' To make this more precise, among bilingual children, those that have more experience with English than Spanish should perform better on English non-words than their peers with less English experience. Preliminary results of an ongoing meta-analysis suggest significant associations between exposure to a given language and performance in both language-dependent and language-independent NWR (Farabolini, Taboh, Ceravolo, \& Guerra, 2021). In any case, this line of research focuses on links between exposure to a given language and NWR performance. In contrast, when we discuss linguistic or cultural differences here, we ask the question of whether children vary in their performance as a function of which language they are learning (e.g., the language's typological properties) and/or their overall, absolute levels of language experience (not relative levels in a multilingual setting).}

@@

Most Yélî parents are swidden horticulturalists, who occasionally fish. Within a group of households, it is often the case that older adolescents and adults spend their day tending to their farm plots (which may not be nearby), bringing up water from the river, washing clothes, preparing food, and engaging in other such activities. Starting around age two years, children more often spend large swaths of their day playing, swimming, and foraging for fruit, nuts, and shellfish in large (\textasciitilde10 members) independent and mixed-age child play groups (Brown \& Casillas, in press; Casillas et al., 2021). Formal education is a priority for Yélî families, and many young parents have themselves pursued additional education beyond what is locally available (Casillas et al., 2021). Local schools are well out of walking distance for many children (i.e., more than 1 hour on foot or by canoe each day), so it is very common for households situated close to a school to host their school-aged relatives during the weekdays for long segments of the school year. Children start school often at around age seven, although the precise age depends on the child's readiness, as judged by their teacher.

@@

We add some observations that will help us integrate this study into the broader investigation of NWR across cultures. As mentioned previously, there is one report of relatively low NWR scores among the Tsimane', which the authors of that paper interpret as consistent with long-term effects of low levels of infant-directed speech (Cristia et al., 2020). However, Cristia et al. (2020) also point out that this is based on between-paper comparisons, and thus methods and myriad other factors have not been controlled for. The Yélî community can help us gain new insights into this matter because direct speech to children under 3 years is comparably infrequent in this community (in fact it may be infrequent in many settings, including urban ones Bunce et al., under review). Our sample also shares other societal characteristics wih the Tsimane' (e.g., the community is rural and relies on farming, children grow up in wide familial networks, Casillas et al., 2021). Although infant-directed speech has been measured in different ways among the Tsimane' and the Yélî communities, our most comparable estimates at present suggest that Tsimane' young children are spoken to about 4.2 minutes per hour (Scaff, Stieglitz, Casillas, \& Cristia, under review), and Yélî children about 3.6 minutes per hour (Casillas et al., 2021). Thus, if these input quantities in early childhood relate to lower NWR scores later in life, we should observe similarly low NWR scores here as in Cristia et al. (2020).

@@

Many NWR studies are based on a fixed list of 12-16 items that vary in length between 1 and 4 syllables, often additionally varying syllable complexity and/or cluster presence and complexity, and always meeting the condition that they do not mean anything in the target language (e.g., Balladares et al., 2016; Wilsenach, 2013). We kept the same variation in item length and requirement for not being meaningful in the language, but we did not vary syllable complexity or clusters because these are vanishingly rare in Yélî Dnye. We also increased the number of items an individual child would be tested on, such that a child would get up to 23 items to repeat (other work has also used up to 24-46 items: Jaber-Awida, 2018; Kalnak et al., 2014; Piazzalunga et al., 2019), with the entire test inventory of 40 final items distributed across children. We used a relatively large number of items to explore correlations with length and phonological complexity. However, aware that this large item inventory might render the task longer and more tiresome, we split items across children. Naturally, designing the task in this way may make the study of individual variation within the population more difficult because different children are exposed to different items.

@@

Additionally, we estimated the usage frequency of the phones present in the target items in a corpus of child-centered recordings (Casillas et al., 2021). That corpus was constituted by sampling from audio-recordings (7--9 hours long), collected as 10 children aged between 1 month and 3 years went about their day. The researchers selected 9 2.5-minute clips randomly and 11 1- or 5-minute clips by hand (selected to represent peak turn-taking and child vocal activity). These clips were segmented and transcribed by the lead researcher and a highly knowledgeable local assistant, who speaks Yélî Dnye natively, has ample experience in this kind of research, and often knew all the recorded people personally. For more details, please refer to Casillas et al. (2021).

@@

Previous work typically reports two scores: a binary word-level exact repetition score, and a phoneme-level score, defined as the number of phonemes that can be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt; as in Cristia et al., 2020). Previous work does not use distance metrics, but we report these rather than the phoneme-level scores because they are more informative. To illustrate these scores, recall our example of an English target being /bilik/ with an imagined response {[}bilig{]}. We would score this response as follows: at the whole item level this production would receive a score of zero (because the repetition is not exact); at the phoneme level this production would receive a score of 80\% (4 out of 5 phonemes repeated exactly); and the phone-based Levenshtein distance for this production is 20\% (because 20\% of phonemes were substituted or deleted). Notice that the phone-based Levenshtein distance is the complement of the phoneme-level NWR score. An advantage of using phone-based Levenshtein distance is that it is scored automatically with a script, and it can then easily be split in terms of deletions and substitutions (insertions were not attested in this study).

@@

There are some reasons to believe that Yélî Dnye put that hypothesis to a critical test: The phoneme inventory is both large and acoustically packed, in addition to containing several typologically infrequent (or unique) contrasts. One could then predict that correlations with typological frequency should be relatively weak because the ambient language puts more pressure on Yélî children to distinguish (perceptually and articulatorily) fine-grained phonetic differences than what is required of child speakers of other languages. On the other hand, it is also possible that this pressure gives Yélî children no benefit, and that some of these categories are simply acquired later in development. We can draw a parallel with children learning another Papuan language, Ku Waru, which has a packed inventory of lateral consonants; children do not produce adult-like realizations of the more complex of these laterals (the pre-stopped velar lateral /ɡ͡ʟ/) until 5 or 6 years of age (Rumsey, 2017).

@@

We investigated the effect of item complexity on NWR scores by varying the number of syllables in the item. In broad terms, children should have higher NWR scores for shorter items. That said, previous work summarized in the Introduction has shown both very small (e.g., Piazzalunga et al., 2019) and very large (e.g., Cristia et al., 2020) effects of stimulus length. Setting aside our monosyllabic stimuli (which contained typologically infrequent segments with lower NWR scores, as just discussed), we examined effects of item length among the remaining stimuli, which range between 2 and 4 syllables long. The effect of item length was not significant in a statistical model that additionally accounted for age and random effects of item and participant. We do not have a good explanation for why samples in the literature vary so much in terms of the size of length effects, but two possibilities are that this is not truly a length effect but a confound with some other aspect of the stimuli, or that there is variation in phonological representations that is poorly understood. We explain each idea in turn.

@@

In broad terms, we expected that NWR scores would increase with participant age, as this is the pattern observed in several previous studies (English Vance et al., 2005; Italian Piazzalunga et al., 2019; Cantonese Stokes et al., 2006; but not in Cristia et al., 2020). Indeed, age was significantly correlated with NWR scores and it also showed up as a significant predictor of NWR score when included as a control factor in the analyses of both item length and average segmental frequency. In brief, our results underscore the idea that phonological development continues well past the first few years of life, extending into middle childhood and perhaps later (Hazan \& Barrett, 2000; Rumsey, 2017).

In contrast, previous work varies with respect to correlations of NWR scores with maternal education (e.g., Farmani et al., 2018; Kalnak et al., 2014; Meir \& Armon-Lotem, 2017). We did not expect large correlations with maternal education in our sample for two reasons: First, education on Rossel Island is generally highly valued and so widespread that little variation is seen there; second, formal education is not at all essential to ensuring one's success in society and may not be a reliable index of local socioeconomic variation locally. In fact, maternal education correlated with NWR score at about r\textasciitilde.1, which is small. We find correlations of about that size for participant sex, which is aligned with previous work (Chiat \& Roy, 2007).

@@

The fourth research area to which we wanted to contribute pertained to the use of NWR across languages and populations, since when designing this study we wondered whether NWR was a \emph{culture-fair} test of phonological development. Although our data cannot answer this question because we have only sampled one language and population here, we would like to spend some time discussing the integration of these results to the wider NWR literature. It is important to note at the outset that we cannot obtain a final answer because integration across studies implies not only variation in languages and child-rearing settings, but also in methodological aspects including non-word length, non-word design (e.g., the syllable and phone complexity included in the items), and task administration, among others. Nonetheless, we feel the NWR task is prevalent enough to warrant discussion about this, similarly to other tasks sometimes used to describe and compare children's language skills across populations, like the recent re-use of the MacArthur-Bates Communicative Development Inventory to look at vocabulary acquisition across multiple languages (Frank et al., 2017).
